Rule-Based Normalization of German Twitter Messages

### 2013-08-14 darmstadt_gscl_2013/article_twitter_normalization.tex
final sections updated with stubs

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{makeidx}  % allows for indexgeneration
\usepackage{paralist} % for inparaenum
\usepackage{multirow} % for multirow
\usepackage{tabularx} % for centering in table
\usepackage{url}

@@

\subsection{Replacement of Twitter-Specific Phenomena}
According to Parker (\cite{parker}), hasthags were supposedly
introduced by Chris Messina, a renowned advocate of open-source
community, in August 2007. The ``hash godfather'', as Messina
describes himself, presumably borrowed the idea of marking relevant
topics with the ``\#''- sign from Internet Relay Chats (IRC) -- the
forebears of modern social networks -- which have been using the pound
character for marking their channel names since the early 1990s.

@@

In our system, unambiguous tokens like, for example, e-mail addresses
were replaced with an artificial word ``\%Mail''. Emoticons were
substituted by tokens ``\%PosSmiley'' or ``\%NegSmiley'', depending on
the type of emotion presumably conveyed by these expressions. In cases
when an emoticon was ambiguous with regard to its polarity, it was
replaced with the special word ``\%Smiley''. Furthermore, leading
``\#{}'' characters were stripped off from the beginning of hashtags
thus leaving only the alphabetic part of these tokens.

@@

A closer look at unknown words which were classified as misspellings
revealed that a prevailing majority of them was considered as
colloquial spelling variants. Such spellings accounted for 70.5~\% of
detected spelling mistakes in \texttt{hunspell} data and 66.17~\% of
misspellings recognized by \texttt{TreeTagger}. Frequency distribution
of colloquial and non-colloquial misspellings for data analyzed with
either tools is shown in Figures \ref{fig:misspl-hunspell} and
\ref{fig:misspl-ttagger}.
...
%% \begin{minipage}[t]{0.45\textwidth}
%%   \begin{figure}[!ht]\label{fig:misspl-hunspell}
%%   \end{figure}
%% \end{minipage}

%% \begin{minipage}[t]{0.45\textwidth}
%%   \begin{figure}[!ht]\label{fig:misspl-ttagger}
%%   \end{figure}
%% \end{minipage}

As can be seen from the Figures, both kinds of spelling variations
have nearly Zipfian distribution. However colloquial spelling
variants, if used, tend to be used frequently whereas non-colloquial
misspellings are rather represented by sporadic singleton cases. This
could be explained by fact that colloquial spellings are usually
formed by systematic rewriting processes. These processes are supposed
to make written form of words more similar to their casual soundings
in everyday speech. According to our data, the most productive of such
processes were:
\begin{itemize}
  \item Omissions of `e' in unstressed (``schwa'') positions,
    e.g. \textit{w\"urd}, \textit{zuguckn} etc. In cases when `e' was
    part of the impersonal pronoun ``es'' and followed a verb,
    remaining `s' was usually appended to the preceding verb form;

  \item Complete omissions or replacement of final consonants with
    their voiceless equivalents, e.g. \textit{nich} instead of
    \textit{nicht} or \textit{Tach} instead of \textit{Tag};

@@

%
\section{Evaluation}\label{evaluation}
To evaluate our system, we used metrics

@@

%% This article provided an overview of exisiting approaches to noisy
%% text normalization task. Additionally, all mentioned methods were
%% classified on the basis of two independent criteria. In section
%% \ref{error:analysis}, we performed qulitative and quantitive analyses
%% of out-of-vocabulary words in German tweets and suggested a set of
%% ad-hoc techniques for mitigating their potential negative influence on
%% natural language processing. This procedure allowed us to reduce the
%% total OOV rate by ... \% for \texttt{hunspell} and by ... \% for
%% \texttt{TreeTagger}.

%% Nevertheless, we should honestly admit that our system still has
%% potential for development and research, since it mainly addresses only
%% one of three main groups of OOV tokens. Future directions should
%% certainly include a more thorough tackling of unintentional spelling
%% mistakes and especially their most prominent types -- deletions and
%% substitutions.

%% Furthemore, a better evalution technique as well as comparison with
%% other systems are needed for our normalization procedure. On the one
%% hand, an \textit{extrinsic} evaluation should be performed (cf. Sparck
%% Jones and Galliers, \cite{sparck}) which means that we not only have
%% to show how the rate of OOV words goes down but much more how this
%% lower OOV rate affects the work of the whole NLP system. On the other
%% hand, we need to assess the quality of our procedure's work on the
%% basis of metrics used by other researchers.

%% One possible estimation criterion which was used by Aw et
%% al. (\cite{aw}), Kaufmann (\cite{kaufmann}), Beaufort et
%% al. (\cite{beaufort}), and Oliva et al. (\cite{oliva}) is the BLEU
%% score (Papineni et al., \cite{papineni}). Another possibility would be
%% to use the Word (WER) and Sentence Error Rates (SER) as suggested by
%% Kobus et al. (\cite{kobus}). However, an obvious difficulty that we
%% already encountered here is that both metrics highly rely on a
%% subjective notion of the look of a ``normalized'' message. While the
%% BLEU score could be an approriate criterion for normalization of
%% messages like ``i luv ma mather and wd do evrythin 4 her'' which in
%% fact looks like as if it were not English. But such highly distorted
%% tweets are rather atypical for German. In this regard, WER and SER
%% could be considered as more appropriate measurement criteria. But
%% these metrics once again are rather dealing with spelling mistakes and
%% would highly depend on whether one, for example, would consider the
%% hash sign in a hashtag as an error.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{makeidx}  % allows for indexgeneration
\usepackage{paralist} % for inparaenum
\usepackage{multirow} % for multirow
\usepackage{tabularx} % for centering in table
\usepackage{graphicx} % for \includegraphics
\usepackage{url}

@@

\subsection{Replacement of Twitter-Specific Phenomena}
According to Parker (\cite{parker}), hasthags were presumably introduced by
Chris Messina, a renowned advocate of open-source community, in August
2007. The ``hash godfather'', as Messina describes himself, allegedly borrowed
the idea of marking relevant topics with the ``\#''- sign from Internet Relay
Chats (IRC) -- the forebears of modern social networks -- which have been
using the pound character for marking their channel names since the early
1990s.

@@

In our system, unambiguous tokens like, for example, e-mail addresses were
replaced with an artificial word ``\%Mail''. Emoticons were substituted by
tokens ``\%PosSmiley'' or ``\%NegSmiley'' depending on the type of emotion
presumably conveyed by these expressions. In cases when an emoticon was
ambiguous with regard to its polarity, it was replaced with the special word
``\%Smiley''. Furthermore, leading ``\#{}'' characters were stripped off from
the beginning of hashtags thus leaving only the alphabetic part of these
tokens.

@@

A closer look at unknown words which were classified as misspellings revealed
that a prevailing majority of them was considered as colloquial spelling
variants. Such variants accounted for 70.5~\% of detected spelling mistakes in
\texttt{hunspell} data and for 66.17~\% of misspellings recognized by
\texttt{TreeTagger}. Frequency distributions of colloquial and non-colloquial
misspellings in data analyzed with either tools are shown in Figures
\ref{fig:misspl-hunspell} and \ref{fig:misspl-ttagger}.
\begin{figure}[ht]
  \centering
  \begin{minipage}[b]{0.45\linewidth}
    \caption{Frequency distributions of non-standard spellings in data
      analyzed with \texttt{hunspell}}
    \includegraphics[width=\textwidth]{data/img/hunspell_spell.png}
    \label{fig:misspl-hunspell}
  \end{minipage}
  \hspace{0.5cm}
  \begin{minipage}[b]{0.45\linewidth}
    \caption{Frequency distributions of non-standard spellings in data
      analyzed with \texttt{TreeTagger}}
    \includegraphics[width=\textwidth]{data/img/treetagger_spell.png}
    \label{fig:misspl-ttagger}
  \end{minipage}
\end{figure}

As can be seen from the figures, both kinds of spelling variations have nearly
Zipfian distribution. However colloquial spelling variants, if used, tend to
be used frequently whereas non-colloquial misspellings are rather represented
by sporadic singletons. This could be explained by the fact that colloquial
spellings are usually formed by systematic rewriting processes. These
processes are supposed to make written form of words more similar to their
casual soundings in everyday speech. According to our data, the most
productive of such processes were:
\begin{itemize}
  \item Omissions of `e' in unstressed (``schwa'') positions,
    e.g. \textit{w\"urd}, \textit{zuguckn} etc. In cases when `e' was part of
    the impersonal pronoun ``es'' following a verb, remaining `s' of this
    pronoun was usually appended to the preceding verb form;

  \item Complete omissions or replacements of final consonants with their
    voiceless equivalents, e.g. \textit{nich} instead of \textit{nicht} or
    \textit{Tach} instead of \textit{Tag};

@@

Since all of the above processes followed some specific formation patterns. We
developed a set of reverese transformation rules which first captured tokens
with suspicious character sequences, checked them in the dictionary, and if
these words were not found there, a transformation associated with each
matched incorrect character sequence was applied.

TODO: more detailed and coherent description of rule module

Unfortunately, spelling variants which were classified as true misspellings
did not show any regularities except for cases of incorrect spelling of sharp
``s'' and umlauts. While writting of ``ss'' instead of sharp ``s'' could also
be considered as correct variant with regard to the Swiss norm. For umlauts, a
straightforward replacement of character sequences like ``ae'', ``oe'' or
``ue'' because these character sequences also appeared in regular German words
like ``israelisch'' or ``quer''. Therefore a set of exceptional cases had to
be found before applying substitutions. As a possible source of such
exceptions we considered words classified by us as in-vocabulary or valid
out-of-vocabulary terms. Another source of possible exceptions was a corpus of
newspaper texts which by definition were supposed to contain only proof-read
content. To derive exceptional character sequences, we subsequently expanded
the substrings to the left and to the right or simultaneously to both
directions until these subsequence only captured words present in our
``normal'' data without capturing any misspeling from the list. After having
collected exceptions, we only applied replacement operation to text spans
lying in between exceptional substrings. On another test set of 500 cases,
this method turned out to have made only one mistake. It replaced the word
``zuende'' with ``z\"unde'' in the sentence ``RT @Netznarr: Als Amtstr\"ager
wei\ss man, dass es zuende geht, wenn man in Umfragen unbeliebter als
Westerwelle ist. \#Wulff''

TODO: should be described as algorithm, should the same approach be tried for
other misspelling char sequences?

%
\section{Evaluation}\label{evaluation}
To evaluate our system, we first performed an intrinsic evaluation by
measuring the rate by which the amount of unknown tokens was reduced for
\texttt{TreeTagger} and \texttt{hunspell}.

TODO: figures for OOV rates after normalization.

To comply with metrics commonly used for evaluating text normalization
procedures for other languages, we also measured word error rate (WER) and
sentence error rate (SER) for restoration of non-standard and incorrect
spellings.

TODO: WER and SER figures after normalization.

Finally, as a first step towards an extrinsic evaluation, we measured
\texttt{TreeTagger}

TODO: Figures for tagging quality before and after normalization.

@@

With this article, we hope to have provided a better insight into the nature
and composition of ill-formed words in German Twitter messages. As was shown
in Section \ref{error:analysis}, special markup elements and casual sppellings
account for more than a half unknown words discovered in Tweets. Furthermore,
almost three quarters of non-standard spellings could be regarded as
colloquial spelling variants rather than occasional slip of the finger. Such
colloquial spellings also showed the tendency to be formed by well formalized
processes and to be used over and over again in texts.

We also suggested a rule-based text normalization approach which could serve
as a baseline comparison measure for future normalization methods which may be
suggested for German tweets. As was shown in previous sections, our approach
could effectively address some ofthe most frequent phenomena which contribute
to a higher rate of out-of-vocabulary words in Twitter texts such as
Twitter-specific elements and non-standard spellings.

We are going to make our classification and test data available online at
\url{http://dasom.ling.uni-potsdam.de/} under the terms consistent with
Twitter regulations. Our classified data could, in our opinion, significantly
save tedious manual work for other researchers.

Admittedly, our method still could be further refined and improved. As
possible steps for such refinement, we see a possible addition of a
machine-learning classifier which could help distinguish spelling mistakes
from valid unknown words. On the other side, a better disambiguation of
suggested replacement variants for misspellings could be achieved by letting
part-of-speech tagger process a word grid with input tokens and their
suggested replacements. At the end of tagging, not only the most probable tag
sequence but the most likely word/tag mesh could be chosen in a single run. At
any rate, these steps are beyond the scope of this work. But we hope to have
laid the groundwork for them.
